### 2023-12-19 supplementary_thev_trxptome.Rmd
wrote some E2 reg results and added gels designed primers for E2 trxpt_15 and pol

---
title: "Supplementary Materials: Elucidating the Transcriptome of Turkey Hemorrhagic Enteritis Virus"
output: pdf_document
geometry: margin=1in
font-size: 11pt
header-includes:
  - \usepackage{helvet}
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left]{lineno}
  - \linenumbers
editor_options:
  chunck_output_type: console
---
\vspace{5mm}

**Running Title:** Novel Insights into Turkey Hemorrhagic Enteritis Virus Transcriptome

\vspace{10mm}
Abraham Quaye^1\*^, Bret Pickett^\*^, Joel S. Griffitts^\*^, Bradford K. Berges^\*^, Brian D. Poole${^\dagger}$^\*^

\vspace{5mm}

^\*^Department of Microbiology and Molecular Biology, Brigham Young University       
^1^First-author      
${^\dagger}$ Corresponding Author

\vspace{5mm}
**Corresponding Author Information**     
[brian_poole@byu.edu](mailto:brian_poole@byu.edu)       
Department of Microbiology and Molecular Biology,       
4007 Life Sciences Building (LSB),    
Brigham Young University,   
Provo, Utah  
\newpage

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
library(knitr)
library(tidyverse)
library(flextable)
source("scripts/r/bam_file_analysis.R")
source("scripts/r/reg_by_reg_plots.R")
```

## Supplementary Table 1A

```{r Sup Table 1A}
flextable(tp_reg_expr_12) %>% theme_zebra()
```

## Supplementary Table 1B

```{r Sup Table 1B}
flextable(tp_reg_expr_24) %>% theme_zebra()
```

## Supplementary Table 1C

```{r Sup Table 1C}
flextable(tp_reg_expr_72) %>% theme_zebra()
```

\newpage
## Supplementary PCR Methods
```{r Supplementary PCR Methods}
set_flextable_defaults(font.size = 3.5)

flextable(supp_pcr_meth_tab) %>%
  theme_zebra() %>%
  flextable::fontsize(j = c("forwardP", "reverseP"), size = 3) %>%
  flextable::align(align = "center", part = "all") %>%
  flextable::width(., width = 0.35) %>% 
  colformat_image(j = "gel_image", i = c(1, 3:7, 9, 10), na_str = "-",
                  width = 1, height = 1) %>%
  flextable::width(j = ~forwardP, width = 1) %>%
  flextable::width(j = ~reverseP, width = 1) %>%
  # flextable::width(j = ~valid_status, width = 0.35) %>%
  flextable::width(j = ~gel_image, width = 1) %>% 
  footnote(., i = c(1:6, 7, 9, 10), j = 12, value = as_paragraph(c("Primer binds exon 1")),
           ref_symbols = c("a"), part = "body") %>%
  footnote(., i = c(1:6, 7, 9, 10), j = 13, value = as_paragraph(c("Primer binds terminal exon")),
           ref_symbols = c("b"), part = "body")
```