### 2024-01-17 supplementary_thev_trxptome.Rmd
added more text to manuscript

---
title: "Supplementary Materials: Elucidating the Transcriptome of Turkey Hemorrhagic Enteritis Virus"
output:
  word_document: default
  pdf_document: default
geometry: margin=1in
font-size: 11pt
header-includes:
  - \usepackage{helvet}
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{textcomp}
editor_options:
  chunck_output_type: console
---
\vspace{5mm}

**Running Title:** Novel Insights into Turkey Hemorrhagic Enteritis Virus Transcriptome

\vspace{10mm}
Abraham Quaye^1\*^, Bret Pickett^\*^, Joel S. Griffitts^\*^, Bradford K. Berges^\*^, Brian D. Poole${^\dagger}$^\*^

\vspace{5mm}

^\*^Department of Microbiology and Molecular Biology, Brigham Young University       
^1^First-author      
${^\dagger}$ Corresponding Author

\vspace{5mm}
**Corresponding Author Information**     
[brian_poole@byu.edu](mailto:brian_poole@byu.edu)       
Department of Microbiology and Molecular Biology,       
4007 Life Sciences Building (LSB),    
Brigham Young University,   
Provo, Utah  
\newpage

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
library(knitr)
library(tidyverse)
library(flextable)
source("scripts/r/bam_file_analysis.R")
source("scripts/r/reg_by_reg_plots.R")
```

## Supplementary Table 1A

```{r Sup Table 1A}
set_flextable_defaults(font.size = 8)
make_reg_stats_table <- function(tp_table){
  return(tp_table %>%
    select(-junc_freq) %>% 
    flextable() %>% 
    theme_zebra() %>% 
    set_header_labels(timepoint = "Time", region = "Region", strand = "Strand",
                      sum_reads_reg = "Total Reads", perc_reg = "Percentage") %>% 
    flextable::width(width = 0.8) %>% 
    flextable::align(align = "center", part = "all") %>%
    flextable::set_table_properties(layout = "fixed", align = "left",
                       opts_pdf = list(caption_repeat = FALSE),
                       opts_word = list(split = F))
  )
}
make_reg_stats_table(tp_reg_expr_12) %>%
  set_caption("Table 1a: Most Transcriptionally Active Regions of THEV at 12h.p.i")

```

## Supplementary Table 1B

```{r Sup Table 1B}
set_flextable_defaults(font.size = 8)
make_reg_stats_table(tp_reg_expr_24) %>%
  set_caption("Table 1b: Most Transcriptionally Active Regions of THEV at 24h.p.i")
```

## Supplementary Table 1C

```{r Sup Table 1C}
set_flextable_defaults(font.size = 8)
make_reg_stats_table(tp_reg_expr_72) %>%
  set_caption("Table 1c: Most Transcriptionally Active Regions of THEV at 72h.p.i")
```

\newpage
## Supplementary PCR Methods
```{r Supplementary PCR Methods}
set_flextable_defaults(font.size = 3.5)

supp_pcr_meth_tab %>% 
  select(-valid_status) %>% 
flextable() %>%
  theme_zebra() %>%
  set_header_labels(trxpt_id = "Transcript ID", region = "Region", num_exons = "Number of Exons",
                    full_trxpt = "Full Transcript", exon1 = "Exon 1", exon2 = "Exon 2",
                    exon3 = "Exon 3", exon4 = "Exon 4", exon5 = "Exon 5", exon6 = "Exon 6",
                    exon7 = "Exon 7", forwardP = "Forward Primer", reverseP = "Reverse Primer",
                    gel_image = "Agarose Gel") %>% 
  flextable::fontsize(j = c("full_trxpt", "exon1", "exon2", "exon3",
                            "exon4", "exon5", "exon6", "exon7",
                            "forwardP", "reverseP"), size = 3) %>%
  flextable::align(align = "center", part = "all") %>%
  flextable::width(., width = 0.35) %>% 
  colformat_image(j = "gel_image", i = c(1:3, 5:25, 27:29),
                  width = 1.1, height = 1) %>%
  flextable::width(j = ~forwardP, width = 1.1) %>%
  flextable::width(j = ~reverseP, width = 1.1) %>%
  flextable::width(j = ~gel_image, width = 1.15) %>%
  # flextable::height(height = 0, part = "header") %>% 
  footnote(., i = c(1:17, 18:29), j = 12, value = as_paragraph(c("Primer binds inside first exon")),
           ref_symbols = c("a"), part = "body", inline = T) %>%
  footnote(., i = c(1:7, 9:17, 20:28), j = 13, value = as_paragraph(c("Primer binds inside terminal exon")), ref_symbols = c("b"), part = "body", inline  = T) %>% 
  # transcripts too long to validate with one primer pair
  footnote(., i = c(8,18,19,29), j = 13, value = as_paragraph(c("Primer binds inside fourth exon")),
           ref_symbols = c("c"), part = "body", inline = T) %>% 
  footnote(., i = c(8), j = 14, value = as_paragraph(c("Agarose gel identical to TRXPT_6 due to identical splicing")), ref_symbols = c("I"), part = "body", inline = T) %>%
  footnote(., i = c(9), j = 14, value = as_paragraph(c("Agarose gel identical to last 3 exons of TRXPT_10 due to identical splicing")), ref_symbols = c("II"), part = "body", inline = T) %>%
  footnote(., i = c(11), j = 14, value = as_paragraph(c("Agarose gel identical to last 4 exons of TRXPT_11 due to identical splicing")), ref_symbols = c("III"), part = "body", inline = T) %>% 
  footnote(., i = c(15), j = 14, value = as_paragraph(c("Agarose gel dentical to TRXPT_23 due to identical splicing")), ref_symbols = c("IV"), part = "body", inline = T) %>% 
  footnote(., i = c(18), j = 14, value = as_paragraph(c("Agarose gel identical to TRXPT_9 due to identical splicing")), ref_symbols = c("V"), part = "body", inline = T) %>% 
  footnote(., i = c(19), j = 14, value = as_paragraph(c("Agarose gel identical to TRXPT_14 due to identical splicing")), ref_symbols = c("VI"), part = "body", inline = T) %>%
  flextable::set_table_properties(layout = "fixed", align = "left",
                       opts_pdf = list(caption_repeat = F),
                       opts_word = list(split = F)) %>%
  set_caption("Table 1: Agarose Gels Showing PCR Amplification of THEV cDNA With Gene-Specific Primers")
```

In the table above, the restriction sites in the primer tails are shown in lowercase letters. All the primer melting temperatures (TMs) are 58-60\textdegree C using a hot start Taq DNA polymerase. The PCR reaction mix was done per manufacturer's instructions. The PCR cycling conditions were as follows: Initial denaturation -- 95$^{\circ}$C for 1 minute; cyclical denaturation -- 95$^{\circ}$C for 30 seconds, annealing -- variable temperature (at least 53$^{\circ}$C) for 30 seconds, primer extension -- 68$^{\circ}$C for variable time, and final elongation -- 68$^{\circ}$C for 5 minutes. We used 35 cycles of amplification.

## Supplementary Computational Analysis   
Our trimmed RNA-seq reads were mapped to the genome of _M. gallopavo_ (with THEV's genome as one of its chromosomes) using `Hisat2` to generate the alignment (BAM) files and `StringTie` used to assemble the transcriptome with a `GTF` file containing the predicted THEV ORFs as a guide. `StringTie` was also used to estimate the normalized expression levels (FPKM) of all the transcripts and `Ballgown` was used to perform statistical analysis and comparisons of the transcript expression levels, which instructive in understanding the temporal regulation THEV genes. We used `RegTools` to extract and analyze the splice junctions in the BAM files. The command `regtools junctions extract` provides a wealth of information about all the splice sites in the BAM file provided such as: the start and end positions, the strand, and number of reads supporting the splice junctions. The command `regtools junctions annotate` gives even more information such as: the splice site donor-acceptor sequences and transcripts/genes that overlap the junction. These information was the basis for estimating and comparing the splicing activity of different regions (TUs) of THEV over time.

The entire analysis pipeline was managed with `Snakemake`, from which we generate the below diagram of the major steps of the analysis pipeline.  
![](project_map.png)